### 2009-10-02 papers/esop2010/main.tex
go

\documentclass[a4paper]{llncs}

\usepackage[utf8x]{inputenc}
\usepackage[T1]{fontenc} 
\input{utf8symbols}
\usepackage{stmaryrd}
\usepackage{amsmath,amssymb}
\usepackage{mathpartir}

\newcommand{\wpre}{{\bf wp}}
\newcommand{\sep}{ ~|~ }
\newcommand{\letml}{{\bf let}}
\newcommand{\inml}{{\bf in}}
\newcommand{\ifml}{{\bf if}}
\newcommand{\thenml}{{\bf then}}
\newcommand{\elseml}{{\bf else}}
\newcommand{\refml}{{\bf ref}}
\newcommand{\propml}{{\bf prop}}
\newcommand{\boolml}{{\bf bool}}
\newcommand{\recml}{{\bf rec}}
\newcommand{\efft}[1]{\langle #1 \rangle}
\newcommand{\lift}[1]{\langle #1 \rangle}
\newcommand{\alist}[1]{\overline{#1} }
\newcommand{\Trueml}{{\bf True}}
\newcommand{\Falseml}{{\bf False}}
\newcommand{\trueml}{{\bf true}}
\newcommand{\falseml}{{\bf false}}
\newcommand{\unitml}{{\bf unit}}
\newcommand{\coq}{Coq}
\newcommand{\who}{Who}
\newcommand{\ocaml}{Ocaml}
\newcommand{\correct}[1]{c(#1)}

\newcommand{\ceil}[1]{\lceil #1 \rceil}

\begin{document}
\title{\mbox{Specifying Imperative Higher-Order Programs}}

\author{Johannes Kanig \and Jean-Christophe Fill\^atre}
\institute{
Laboratoire de Recherche en Informatique, Univ Paris-Sud, CNRS, Orsay F-91405\\
INRIA Futurs, ProVal, Parc Orsay Universit\'e, F-91893
}
%TODO affiliation
%TODO explain somewhere the correspondence store type <-> map, effect variable
%<-> domain, region <-> key
%TODO the constants dealing with maps

%external TODO : proofs for Hashtbl
%external TODO : proofs of system
%

\maketitle

\begin{abstract} We present a specification language suited for reasoning
  about effectful higher-order programs. The language's main features are
  higher-order logic and effect polymorphism, both of which enable a truly
  modular treatment of higher-order functions.  We also define a weakest
  precondition calculus for this language which generates proof obligations in
  higher-order logic.  To our knowledge, this is the first time that effect
  polymorphism is put to use in program verification.  A simple mechanism
  based on regions permits to restrict the analysis to alias-free programs,
  thus obtaining much simpler proof obligations. Several examples demonstrate
  the practicality of the system. A prototype implementation exists and has
  been used to prove the correctness of these and more examples.
\end{abstract}

\section{Introduction}

The purpose of the present paper is to introduce a simple weakest precondition
calculus for imperative higher-order programs. Higher-order functions are a
key feature, not only of purely functional languages like Haskell and the pure
fragment of ML, but also in combination with imperative features, like
references. However, reasoning about higher-order functions is difficult, and
especially so in the presence of effectful computations. Any function's
specification depends on its arguments. In a higher-order setting, arguments
may be functions that have their own specification. In consequence, to be
modular, the specification of a higher-order function can only be formulated
{\em depending on the specifications of its functional arguments}. In a
setting where side effects are possible, one also needs to formulate the
specification {\em depending on the side effects of the functional arguments}.
We claim that to be truly modular, the specification of a higher-order
function must be generic with respect to the specification {\em and} the
effect of its arguments.

In this paper, we first present a specification language combining features of
a programming language and logic. It uses an effect system and a region
calculus to track references. The effect system contains {\em effect
polymorphism}, thus dealing well with higher-order functions. The
specifications can be expressed in higher-order logic, to be able to state
properties of functional arguments. We show how two very simple modifications
to the system let us obtain alias-free programs, with the aim of obtaining
much simpler proof obligations.

We then proceed to define a weakest precondition calculus based on this
language, transforming an annotated program into a logic formula, which has to
be valid for the program to be correct. This calculus is very straight-forward
and syntax-directed.

Finally, we show the usefulness of the system by giving several examples of
annotated programs. All the programs have been proved correct using our
prototype implementation called \who.

\paragraph{Outline.}
Section~\ref{sec:genpic} gives an introduction to the described language as
well as an example. Sections~\ref{sec:syntax} and \ref{sec:wp} give the
details about the programming language and the weakest precondition calculus.
In section~\ref{sec:implem} we speak about the implementation and show some
examples. Finally, section~\ref{sec:related} discusses related work.

\section{The general picture and a motivating example}
\label{sec:genpic}
%TODO Hoare logic = sharing between logic and prog language + alias free
also one or two examples in the described language

characteristics : 
\begin{itemize}
  \item Hoare logic
  \item logic and program share values  
  \item Type and effect system
  \item Region typing to track references
  \item the wp needs absence of aliasing to produce good POs
\end{itemize}

\section{A programming language for specification}
\label{sec:syntax}

technical presentation:
\begin{itemize}
  \item  notion of substitution - correspondence χ κ
  \item semantics is standard (goes to appendix )
\end{itemize}

\subsection{Syntax}
\begin{figure}[htpb]
\begin{eqnarray*}
  x & & \text{Program Variables}\\
  α & & \text{Type Variables}\\
  e & & \text{Effect Variables}\\
  ρ & & \text{Region Variables}\\
  χ &::=& α \sep e \sep ρ \\
  t,p,q &::=& x~[\alist{κ}] \sep λx:τ.t \sep \recml~f~(x:τ).\{p\}t\{q\} \sep t~t\sep \\ 
  & & \letml~ x~[\alist{χ}] = t~\inml~t \sep \ifml~t~\thenml~t~\elseml~t\\
  τ &::=& α \sep τ -> τ \sep τ ->^ε τ \sep τ \times τ \sep τ~\refml_ρ \sep
  \efft{ε} \\
  ε &::=& (ρ \sep e)*\\
  κ &::=& τ \sep ε \sep ρ
\end{eqnarray*}
  \caption{Syntax}
  \label{fig:syntax}
\end{figure}

Our programming language (Fig.~\ref{fig:syntax}) is basically the λ-calculus
with \letml-bindings, where there are two different λ-bindings. The first one
($λx:τ.t$) describes a pure, non-recursive binding, and the second one
($\recml~f~(x:τ).\{p\}t\{q\}$) describes recursive, potentially impure
functions; $p$ and $q$ in this term are {\em specifications}, which are
of the same syntax as programs, but are differently typed (typing will be
discussed in the next section). The term $p$ corresponds to the {\em
precondition}, $q$ to the {\em postcondition} of the function. The analogy to
Hoare triples is of course obvious. As usual, there are also variables,
\letml-bindings and an \ifml-\thenml-\elseml-construct. Variables on the type
level can be generalized at \letml-boundaries and have to be instantiated when
the corresponding program variable is used.

Before we describe the syntax of types, let us first explain {\em regions} and
{\em effects}. Regions are simply type level names for memory locations. An
effect is a set of memory locations and {\em effect variables}. An effect
variable stands for an unknown effect. Effects are used to specify
of memory locations an expression can write. They are also used in function
types to describe the memory locations which may be written if the function is
called.

Type, region and effect variables must be explicitly generalized and
instantiated. For this purpose, we have defined the metavariable $χ$,
representing variables, and $κ$, representing the corresponding instantiation
objects types, regions and effects\footnote{We do not distinguish between
region variables and region constants; a region constant is simply a region
variable whose scope is the entire program.}. Now, it can be seen that
generalization takes place at \letml-bindings and instantiation at variable
occurrences. We use a horizontal bar over terms (like in ${\overline{χ}}$) to
denote lists.

\subsection{Typing}

We have already seen that programs and specification share in fact the same
syntax. Though they are differently typed, their typing relations still share
a lot. For this reason, and to underline the fact that in Hoare logic, values
are shared between programs and the logic, we present three different typing
relations. One for programs, one for specifications, and the common part is
factored out in a third typing relation for values. All typing relations
contain an environment $Γ$, which is simply a map from variable names to {\em
type schemes} of the form $∀\alist{χ}.τ$ .

\begin{figure}[htbp]
  \begin{mathpar}
  { \inferrule*[Left=Var] {Γ(x) = ∀\alist{χ}.τ} {Γ|-_v x~[\alist{κ}] :
  τ[\alist{χ}|->\alist{κ}]} } \and
  { \inferrule*[Left=PureFun] {Γ,x : τ' |-_v t : τ} {Γ|-_v λx : τ'.t : τ' ->
  τ} } \and
  { \inferrule*[Left=Rec] 
  { Γ' = Γ, x : τ' \\ Γ'' = Γ',~f : τ' ->^ε τ \\ Γ'' |- t : τ,ε \\ Γ' |-_l
  p : \efft{ε} -> \propml \\ Γ' |-_l q : \efft{ε} ->
  \efft{ε} -> τ -> \propml } 
      {Γ|-_v \recml~f~(x : τ').\{p\}t\{q\} : τ' ->^ε τ} 
  } \and
  { \inferrule*[Left=PureApp]
    {Γ |-_v t1 : τ' -> τ \\ Γ|-_v t2 : τ'}
    {Γ |-_v  t1 ~ t2 : τ}
  }
  \end{mathpar}
  \caption{Value typing}
  \label{fig:valuetyping}
\end{figure}

\paragraph{Value typing.}
Let us first look at the common part concerning values. We define a relation
$Γ|-_v t : τ$, which means ``$t$ is well-typed wrt. the environment $Γ$, and
is of type $τ$''. Values cannot have any effect.  Variables and pure
abstractions are typed in the usual way; type instantiations at variable
occurences are substituted in the type. The rule {\sc REC} simply states that
the body of the function can use the argument $x$ as well as the function $f$.
The specifications $p$ and $q$, however, cannot use $f$ and must be well-typed
wrt. the typing relation $|-_l$ described later on. Finally, pure applications
between values are considered as values as well, as they don't have any
effect.

\begin{figure}[htpb]
\begin{mathpar}
  { \inferrule*[Left=Value]
  { Γ |-_v t : τ  }
  { Γ |- t : τ,\emptyset  }
  } \and
  { \inferrule*[Left=App]
    {Γ |-_v t1 : τ' ->^ε τ \\ Γ|-_v t2 : τ'}
    {Γ |- t1 ~ t2 : τ, ε}
  } \and
  { \inferrule*[Left=ite]
    { Γ |- _v t1 : \boolml \\ Γ |- t2 : τ, ε1 \\ Γ |- t3 : τ, ε2 }
    { Γ |- \ifml~ t1 ~\thenml~ t2 ~\elseml ~ t3 : τ, ε1  ε2 }
  } \and
  { \inferrule*[Left=Letv]
    { Γ, \alist{χ} |-_v t1 : τ' \\ Γ, x : ∀\alist{χ}.τ' |- t2 : τ,ε }
    { Γ |- \letml~ x~[\alist{χ}] = t1 ~ \inml~ t2 : τ, ε}
  } \and
  { \inferrule*[Left=Let]
    { Γ |- t1 : τ', ε \\ Γ, x : τ' |- t2 : τ, ε }
    { Γ |- \letml~ x = t1 ~ \inml~ t2 : τ, ε }
  } \and
  { \inferrule*[Left=Sub]
    { Γ |- t : τ', ε \\ ε \subseteq ε' }
    { Γ |- t : τ, ε' }
  }
\end{mathpar}
  \caption{Program typing}
  \label{fig:progtyping}
\end{figure}

\paragraph{Program typing.}
The typing of programs, written $Γ|- t : τ,ε$, assigns a type $τ$ and an
effect $ε$ to an expression $t$, given an environment. This typing is
presented in {\em A-normal form}, a normal form where all intermediate values
are \letml-introduced. This is not a restriction, as any program can be
transformed an A-normal form.

All values are also programs with no effect (rule {\sc
Value}). An application must consist of two values and the effect of the
overall expression is precisely the effect of the called function (rule {\sc
App}). There are two rules for \letml, due to the value restriction. One only
can generalize type, effect and region variables when the bound expression is
a value. In this case (rule {\sc LetV}), only the second term may produe
effects. In the other case (rule {\sc Let}), no generalization takes place,
but both terms can have effects. The {\sc Let} rule actually obliges both
terms to have the same effect, but together with the {\sc Sub} rule, which can
increase the effect of an expression, this can always be achieved.
Finally, \ifml-expressions are typed in the usual way, both branches have to
have the same effect.

\paragraph{Logic typing.}

The typing of annotations is basically an extension of the typing for values;
logic terms never have any effect. To ensure this, we need to forbid calling
effectful functions in the logic. One way is to simply omit the typing rule
for effectful function calls. There is, however, another way which also
enables us to inspect the pre/ and postcondition of a function. To this end,
let us introduce the operation of {\em lifting} a type to the logic level (see
Fig.~\ref{fig:typelift}).
\begin{figure}[htpb]
  \begin{eqnarray*}
    \ceil{α} &=& α\\
    \ceil{τ~\refml_ρ} &=& \ceil{τ}~\refml_ρ\\
    \ceil{τ -> τ'} &=& \ceil{τ} -> \ceil{τ'}\\
    \ceil{τ \times τ'} &=& \ceil{τ} \times \ceil{τ'}\\
    \ceil{\efft{ε}} &=& \efft{ε}\\
    \ceil{τ ->^ε τ'} &=& (\ceil{τ} -> \efft{ε} -> \propml)\times (\ceil{τ} ->
    \efft{ε} -> \efft{ε} -> \ceil{τ'} -> \propml)
  \end{eqnarray*}
  \caption{Lifting types}
  \label{fig:typelift}
\end{figure}
This translation traverses the type and replaces all effectful types by a
tuple type which describes a tuple of a pre- and a postcondition. The
precondition takes as an argument the argument of the function as well as the
part of the store, at the moment of the function call, which corresponds to
the effect of the function. The postcondition's two first arguments are the
same, but in addition it expects the store {\em at the exit of the function},
as well as the function's return value.

\begin{figure}[htbp]
  \begin{mathpar}
  { \inferrule*[Left=Logic]
    { Γ |-_v t : τ  }
    { Γ |-_l t : \ceil{τ} }
  } \and
  { \inferrule*[Left=LetLogic]
    { Γ, \alist{χ} |-_l : t' : τ' \\ Γ, x: ∀\alist{χ}.τ' |-_l t : τ  }
    { Γ |-_l \letml~x~[\alist{χ}] = t'~\inml~t : τ }
  }\and
  { \inferrule*[Left=IteLogic]
    { Γ |- _l t1 : \boolml \\ Γ |-_l t2 : τ \\ Γ |-_l t3 : τ }
    { Γ |- \ifml~ t1 ~\thenml~ t2 ~\elseml ~ t3 : τ }
  }
  \end{mathpar}
  \caption{Logic typing}
  \label{fig:logictyping}
\end{figure}

Now we can come back to logic typing. Every value is also a logic term,
but now it is assigned its {\em lifted} type. The two other rules are just 
pure variants of their impure counterparts in the program typing.

\paragraph{Predefined constants.}

To actually be able to write programs in this language, we need many
predefined constants and types. The typing relations already mention two,
\propml\ and \boolml. \propml\ is the logical truth and contains the values
\Trueml\ and \Falseml, while \boolml\ describes boolean (decidable) properties
and contains \trueml\ and \falseml. The distinction between \propml\ and
\boolml\ is the same one as in \coq, for example. Properties expressed in
\propml\ are potentially undecidable and one cannot take decisions by testing
their truth value. On the other hand, \boolml\ talks about decidable
properties and so it makes sense to expect the predicate of the
\ifml-statement to be of type \boolml.

The reader may have wondered why the specification language does not contain
any logical connectives. These can of course be declared independently, for
example the conjuction is just a function: \begin{equation*} /\ :  \propml ->
\propml -> \propml \end{equation*} Actually, when equality is given, all
logical operations (even quantification) can be defined in the
logic~\cite{andrews86}.  In the remainder of the paper, we suppose
predefined the usual logical connectives conjunction, disjunction,
implication and equivalence, as well as universal and existential
quantification. We also assume all the arithmetic operations as well as the
constructor $(t,t)$  and the accessors $fst$ and $snd$ for pairs.

%TODO pre and post
% TODO TODO
%Finally, we introduce two symbols
%\begin{eqnarray*}
%  \oplus &:&∀ e1 e2 . \efft{ e1 } -> \efft{ e2 } -> \efft { e1  e2 }\\
%  |_{ε} &:& ∀ e. \efft{ e ε } -> \efft{ ε }.
%  !! & : & ∀αρ e. α~\refml_ρ -> \efft{ ρ e } -> α
%\end{eqnarray*}
%The first one denotes a combining operation for stores; the resulting store
%contains all the mappings of the second store, as well as all the mappings of
%the first one, as long as they are not present in the second one.
%union of the two arguments, in the sense that its domain is the union of its
%arguments.

\subsection{References}

We haven't mentioned references yet, but all we need to support them is to add
the three usual functions $ref, :=, !$ for reference creation, assignment and
lookup, respectively, to the environment. In our system, these functions have
the following types:
\begin{eqnarray*}
  ref &:& ∀αρ.α ->^ρ α~\refml_ρ\\
  := &:& ∀αρ.α~\refml_ρ ->^\emptyset α ->^ρ \unitml\\
  ! &:& ∀αρ. α~\refml_ρ ->^ρ α
\end{eqnarray*}

\subsection{Alias restriction}

The presented programming language, along with the specifications, can, in
fact be used to prove programs. The weakest precondition calculus presented in
the next section can be used to obtain proof obligations which, when proved,
guarantee the correctness of the program with respect to its specification. On
the other hand, in its presented form, and because of the presence of
aliasing, one would constantly need equalities and inequalities between
reference variables to be able to reason about a program.

\begin{figure}[htpb]
  \begin{mathpar}
    {\inferrule*[Left=Var]
      {Γ(x) = ∀\alist{χ}.τ \\ σ = [\alist{χ}|->\alist{κ}] \\ σ \sim τ }
      {Γ |- x : τσ }
    } \\
    { \inferrule*[Left=LetRef, Right={ρ \textup{fresh}, $ρ\notin τ$}]
      {Γ |-_v t1 : τ' \\ Γ, x : τ'~\refml_ρ |- t2 : τ, ε }
      {Γ |- \letml~x = ref~ t1 ~\inml~ t2 : τ, ε \setminus ρ}
    }
  \end{mathpar}
  \caption{A modified and a new rule for aliasing}
  \label{fig:aliasing}
\end{figure}

A way out is to restrict programs to alias free programs. Fortunately, it is
easy to modify the presented typing rules to render all programs with aliasing
ill-typed (see Fig~\ref{fig:aliasing}). In the following, we will use regions
to identify memory locations entirely; each region shall describe precisely
one memory location. Starting from this assumption, a program can actually
exhibit {\em two} forms of aliasing. In the first one, it may use several
names (region variables) for the same memory location. This can happen when a
function is generalized over, say, two region variables, assuming they are
different, and is then instantiated with the same region variable twice. The
classical example is the following program
%TODO
\begin{who}
  let f [ρ ρ'] (x : int ref(ρ)) (y : int ref(ρ')) =
  { }
  x := !x + 1;
  !y
  { r : r = !!y old }
\end{who}

which is only correct if $ ρ $ and $ ρ' $ are indeed different. To enforce
this, we add a side condition to the instantiation: the substitution $σ$ must
be {\em compatible} with the type $τ$. To understand what compatibility means,
consider an {\em atomic} substitution $[ρ|->ρ']$. It is compatible with any
type which does not contain $ρ'$. Similarly, the substitution $[e|-> ρ1  ρ2
e'$] is compatible with any type which does not contain $ ρ1 , ρ2 $ nor $e'$.
Expressed more generally, an atomic substitution is compatible with a type if
the type does not contain any region or effect variables which are in the
range of the substitution. Finally, a composite substitution is compatible
%TODO atomic and composite
with a type if each atomic substitution is compatible with the result of the
application of the previous substitution. In this way, it becomes impossible
to have different names for a single reference cell.

The second form of aliasing is when we use the same name for {\em different}
memory locations. This can happen when we allow the user to call $ref$ several
times with the same region instantiation. A simple possibility is to attach
the suppress the $ref$ function entirely and to replace it by a \letml\refml\ 
construct which combines name creation and memory allocation. In this way,
there can of course be only one reference per name. On the other hand, because
of the side condition $ρ\notin τ$, the reference cannot escape the scope of
its \letml\refml\ construct, and so, a function cannot return a reference
which has been created in its body. This restriction could be leveraged a bit if
one seperates name creation and allocation, while still maintaining that one
can only allocate a single reference in each region. This is how it is done in
the correctness proof of this restriction. The presented \letml\refml\
construct is of course a special case.
%TODO now, all operations on maps can be statically resolved.

\subsection{Some properties}

%TODO
$Γ|-_v v : τ$ implies $Γ |-_l \ceil{v} : \ceil{τ}$

Subject reduction of our type system is a consequence of subject reduction of
the region calculus of Tofte/Talpin~\cite{tofte97ic}.

\section{A Simple Weakest Preconditions Calculus}
\label{sec:wp}

The aim of the weakest precondition calculus is to obtain proof obligations
from the annotated program. To do this, the calculus starts from a formula
which holds, or is supposed to hold, at the end of the program, and walks
backwards in the program source to obtain the weakest formula which has to
hold at the beginning of the program (hence the name {\em weakest
precondition}.

In the following, we use the metasymbol $v$ to denote well-typed values, i.e.
terms with a typing derivation $Γ|-_v v : τ$. In the definition of the weakest
precondition calculus, we will use the notation $(t : τ, ε)$ to denote that
$t$ has type $τ$ and effect $ε$. We assume that applications of the
subeffecting rule {\sc Sub} rule are explicitely inserted in the term; they
are written $(t:ε<:ε')$.

\begin{figure}[htbp]
  \begin{eqnarray*}
    \ceil{x~[\alist{κ}]} &=& x~[\alist{κ}]\\
    \ceil{λx:τ.t} &=& λx:τ.\ceil{t}\\
    \ceil{\recml~f~(x:τ).\{p\} t \{q\}} &=& 
    (λx:\ceil{τ}.p,λx:\ceil{τ}.q)\\
    \ceil{t_{τ' -> τ}~t'} &=& \ceil{t}~\ceil{t'}
  \end{eqnarray*}
  \caption{Lifting values}
  \label{fig:valuelift}
\end{figure}

First of all, analogous to the lifting of types to the logic level, we define
a lifting of values to the logic level. Similarly, this translation traverses
the value until it encounters an impure function. From the logic point of
view, an impure function is represented by its specification, so the
translation keeps only the pair of pre- and postcondition, both abstracted
over the function argument. The function body is thrown away.

\begin{figure}[htbp]
  \begin{eqnarray*}
    c(x~[\alist{κ}]) &=& \Trueml \\
    c(λx:τ.t) &=& ∀x:\ceil{τ}.c(t) \\
    c(t~t') &=& c(t) /\ c(t') \\
    c(\recml~f~(x:τ).\{p\}(t : τ',ε)\{q\}) &=&
    ∀x:\ceil{τ}.∀f:\ceil{τ->^ετ'}.  f = \ceil{v} => \\& &
    ∀s:\efft{ε}.p~s => \wpre_s(t,q~s) \\
    \wpre_s(v, q) &=& c(v) /\ q~s~\ceil{v}\\
    \wpre_s( t_{τ'->^ε τ} t' ,q) &=& \correct{t} /\ \correct{t'} /\
    fst~\ceil{t}~\ceil{t'}~s /\ \\ 
    & & ∀s':\efft{ε}∀ x:\ceil{τ}.~  snd~\ceil{t}~\ceil{t'}~s~s'~x => q~s'~x\\
    \wpre_s(\letml~x~[\alist{χ}] = v~\inml~t, q) &=&
      ∀\alist{χ}.c(v) /\ \letml~x~[\alist{χ}]=\ceil{v}~\inml~ \wpre(t,q)\\
    \wpre_s(\letml~x = ( t1 : τ,ε) ~\inml~ t2 , q) &=&
    \wpre_s( t1 ,λs':\efft{ ε }.λx:\ceil{τ}.\wpre_{s'}( t2 , q ))\\
    \wpre_s(\ifml~ t1 ~\thenml~ t2 ~\elseml~ t3 , q) &=&
      \ifml~ \ceil{ t1 }~\thenml~ \wpre_s( t2 , q))~\elseml~\wpre_s( t3 , q)) \\
      \wpre_s( (t : ε <: ε'), q) &=& \wpre_{s|_ε}(t, λs':\efft{ε}.q~(s\oplus s'))
  \end{eqnarray*}
  \caption{The weakest precondition calculus}
  \label{fig:wp}
\end{figure}

The weakest precondition calculus itself comes in two parts, one for values
and one for effectful terms. Both are reasonably straightforward. The {\em
correctness} of values, written $c(v)$, ensures that any effectful function
appearing in a value adheres to its specification. To this end, the value is
traversed, quantifying over bound variables and using conjunction for pure
applications. When an effectful function is encountered, one has to prove that
the precondition of the function implies the postcondition, given the function
body.

The part concerning effectful functions, denoted $\wpre_s(t,q)$ takes as
argument a term $t$ of type $τ$ and effect $ε$, a postcondition $q$ of type
$\efft{ε} -> τ -> \propml$ and a current state $s$ of type $\efft{ε}$. If the
term is in fact a value, the value has to be correct and the postcondition
must be true for the lifted value and the current state. In the case of an
application, both values must be correct, the precondition of the function
must hold in the current state, and for any state and return value of the
function, the function's postcondition must imply the condition $q$ to be
proved.

As for the typing relation, we have two rules for the \letml-binding. If the
bound term is a value, we can translate the \letml-binding in the program by a
\letml-binding in the language. In the other case, the \wpre\ of $ t2 $ becomes
the argument of the \wpre of $ t1 $ (we denote $ε$ the effect of $ t1 $ and $
t2 $). The rule for \ifml-expressions is straightforward. The rule for
subeffecting adapts the domain of the current store and the expected domain of
the postcondition so that the types work out.

%TODO lemmas about correctness
\section{Implementation}
\label{sec:implem}

A prototype implementation by the name ``\who''
exists~\cite{KanigFilliatre09wml} and is freely available. It implements the
described language with type, region and effect polymorphism, higher-order
logics and store types in the logic. It takes as an input a fully annotated
program and outputs proof obligations which have to be proved to guarantee the
correctness of the program with respect to its specification. Currently, only
output to the interactive proof assistant Coq has been implemented. Our logic
is a subset of Coq's, so the translation is particularly simple.

The implementation consists of less than 3000 lines of \ocaml\ code, including
a limited form of type and effect inference.

\section{Examples}

In this section, we will show two simple examples which use higher-order
features.
\subsection{A simple memoization function}

\begin{figure}[tpb]
\begin{who}
type ('a,'b) map
logic mem : forall ('a, 'b) . 'a -> ('a,'b) map -> bool
logic get : forall ('a, 'b) . 'a -> ('a,'b) map -> 'b
logic set : forall ('a, 'b) . 'a -> 'b -> ('a,'b) map -> ('a,'b) map

let stores f0 m =
  forall x : Z, mem x m = true -> get x m = f0 x

parameter table : (int,int) map ref(ρ)

let memo [e] (f0 : int -> int) (f : int ->{} int) (x : int) = 
  {pre f x /\ stores f0 !!table}
  if mem x !table then get x !table
  else
    let z = f x in
    table := set x z !table;
    z
  {r : post f x r}
\end{who}
  \caption{The code along with the specification of the {\tt memo} function}
  \label{fig:memo}
\end{figure}

Our first example is a simple memoization function {\tt memo} for (pure)
functions.  It takes as an argument a function (from {\tt int} to {\tt int}
here) and an integer and returns the same result as the functional argument
would; however, before executing the argument, it will look up the integer in
a table to see if the function hasn't been called yet on this integer. If so,
the table's value is returned, otherwise we call the function, store the
result in the table, and return it. Fig~\ref{fig:memo} shows the code along
with the specification. The specification is very simple; it states that {\tt
memo} will return a value that verifies the postcondition of {\tt f}, as long
is initially we have the precondition of f and as long as the table stores
only valid key/value pairs. As we are not interested in the exact way of the
implemenation of the map, we simply model it here by a reference to a
functional map with get/set functions.

\subsection{ymemo}

\begin{figure}[tpb]
\begin{who}
  let realizes 
  let ymemo (f0 : int -> int) 
            (ff : (int ->{table} int) ->{} int ->{table} int) = 
  { forall (k : [[int ->{table} int ]]),
      { realizes f0 k} ff k { r : realizes f0 r} }
  let rec f (x : int) : int ->{table} int =
    { stores f0 !!table }
    if mem x !table then get x !table
    else
      let z = ff f x in
      table := set x z !table; z
    {r : r = f0 x /\ stores f0 !!table}
  in f
  { rf : realizes f0 rf }
\end{who}
\caption{The {\tt ymemo function} and its specifacation}
  \label{fig:<+label+>}
\end{figure}<++>

\subsection{All examples - an overview}
\begin{figure}[tbp]
\begin{center}
    \begin{tabular}{ | r | r | r | r |}
    \hline
    Program & Ocaml Code & Code + Spec in \who & Proofs \\ \hline
    Memo/ymemo & 15 loc & 26 loc & 90 lop  \\ \hline
    {\tt Array} & 69 loc & 170 loc  & 320 lop  \\ \hline
    {\tt Hashtbl} (incomplete) & & &  \\ \hline
    Koda-Ruskey~\cite{KanigFilliatre09wml} & 10 loc & 30 loc & $\sim$700 lop  \\ \hline
    \end{tabular}
\end{center}
  \caption{A summary of all programs proved so far using the \who\ tool}
  \label{fig:whoproofs}
\end{figure}

Fig.~\ref{fig:whoproofs} shows an overview over the programs we have been able
to prove using the \who\ tool. The first line summarizes the numbers of the
memo/ymemo example of the previous subsections.

The next two lines concern the two modules {\tt Array} and {\tt Hashtbl} of
the \ocaml\ standard library. For the {\tt Array} module, all functions
present in the \ocaml\ module have been proved, with the exception of the {\tt
Array.map} family, whose specification is very similar to {\tt iter}, and the
{\tt Array.sort} function, which is quite complicated to prove, but not very
interesting concerning higher-order functions or side effects. In this proof
development, arrays are modeled as references to functional arrays with the
usual axioms on access and update.

The proof development concerning the {\tt Hashtbl} module is ongoing work. The
difficult part of this development is the model of the \ocaml\ hash table; In
\ocaml, when one overwrites a previously existing entry in the table, the old
value still exists, and is only {\em shadowed} by the new key/value pair. If
one calls {\tt Hashtbl.remove} using this key, the old value reappears. Also,
if several bindings for a key exist, the key may be encountered several times,
during a call of {\tt Hashtbl.iter}, for example. A possible way of modeling
this is an association list instead of a simple mapping, where keys in front
of the list can shadow the same key further down in the list. 
% TODO Currently, only
%two functions of this module have been proved correct.

Finally, the last line shows the numbers for the proof development of the
Koda-Ruskey algorithm~\cite{KodaRuskey93}. A small, efficient
implementation~\cite{FilliatrePottier02} is shown to be correct. More details
about this development can be found in~\cite{KanigFilliatre09wml}.

\section{Related Work}
\label{sec:related}

%TODO rewrite
%TODO new citation for Ynot
Honda, Yoshida and Berger~\cite{Honda05anobservationally} present a
logic for imperative higher-order programs, using a slightly more
expressive language than ours. They present their calculus using rules
in Hoare logic style, while we presented a \wpre-calculus. Their
system does not provide the same modularity than ours, as one can not
reason modularly about side effects of functional arguments. It also
has the drawback of being difficult to implement, due to the presence
of a large number of non-structural rules. Our system, starting from a
correctly annotated program, is syntax-directed and thus entirely
automated. Their logic of proof obligations is non-standard, while we
obtain obligations in higher-order logic. However, their calculus can
deal easily with functions in references, even if the stored function
reads or modifies the reference it is stored in. While, with the
appropriate extensions, one can reason about functions in references
in our system, this particular kind of programs cannot be reasoned
about. Honda, Yoshida and Berger prove a completeness result for their
system. We have proved no such result. Their extended
system~\cite{BergerHondaYoshida05aliasing} has the same advantages and
drawbacks compared to ours, but additionally deals with aliasing
situations.

The Coq proof assistant~\cite{CoqManualV81}, in particular in
combination with the Program extension~\cite{sozeau07icfp}, can be
used to implement and prove arbitrary properties of purely functional
programs, using dependent types. The appeal of that approach is that
the only limit of expressiveness is the one of Coq. One can also do
meta-reasoning about programs inside Coq and, at the end of the
development, extract a certified implementation in \ocaml\ or
Haskell. The confidence in programs proved in this way is very high,
because in the end one obtains a Coq proof term that corresponds to
the correctness of the program. This term has been checked by Coq's
kernel. However, without extensions such as Ynot (see next paragraph),
it is impossible to implement (and reason about) effectful
computations. Another drawback of this approach is lack of automation.

The Ynot System~\cite{Nanevski08Awkward} is an extension to the Coq
proof assistant, capable of reasoning about imperative higher-order
programs, including effectful functions as arguments, using a monad in
which effectful computations can take place. It uses higher-order
separation logic as the device to speak about modifications of the
state. Ynot uses an (arguably less intuitive) monadic representation
of effectful computations. Ynot is more expressive than our system and
features modular reasoning (abstracting over the specification and
effect of a function as argument) while being able to reason about
aliasing situations thanks to separation logic.  On the other hand,
Ynot shares the Coq's main weakness: lack of automation. While we
haven't yet presented a platform to automatically discharge generated
proof obligations, this seems easily feasible, along the same vein as
Régis-Gianas and Pottier~\cite{regis-gianas-pottier-08}, who argue
that first-order SMT solvers~\cite{RanTin-SMTLIB} are suitable even
for higher-order proof obligations. They also present an encoding
scheme to support their claim. Their practical results are
promising. The Ynot system can not use this kind of automation because
it uses separation logic. Automated provers for separation logic
exist, but we are not aware of any attempts to use these tools in a
higher-order setting.

%TODO also say that an approach like Ynot ou Coq doesn't give any help
%for the mechanical part of the proof.

The Pangolin system, the implementation of the theoretical system
of~\cite{regis-gianas-pottier-08}, can also be used to reason about
purely functional programs. It is also one of the starting points of
our work, and our purely functional fragment is basically the Pangolin
system, although we removed algebraic data types for clarity of
presentation. For this reason, we share the potential advantages (good
automation, relatively simple system) and drawbacks (no
meta-reasoning or observational equality, no machine-checked proof
term). Pangolin cannot deal with side effects.

Another line of work consists in giving a translation from an
imperative to a functional language, with the aim of applying one of
the techniques for purely functional programs to the
translation. O'Hearn and Reynolds~\cite{ohearn-reynolds-00} present a
translation from Algol with first class references and dynamic
allocation to a linear $\lambda$-calculus. They impose an
alias-avoiding restriction which is similar to ours. Charguéraud and
Pottier~\cite{chargueraud-pottier-08} give a translation from Core ML
with first-class references, dynamic allocation and aliasing to a
functional language, using a very expressive type system with linear
capabilities. Both systems handle much more expressive languages than
ours. However, they do not describe a mechanism to obtain proof
obligations, for example a \wpre-calculus. Although we use a simpler
programming language, we present a complete cycle from imperative
programs to proof obligations. We also claim that extending our
language and our \wpre-calculus is simple, as long as aliasing remains
excluded.  We leave these extensions for future work.

Systems for verification of first-order programs like
Why~\cite{Filliatre00a} or Spec\#~\cite{BarnettLS04} are the other
starting points of this work. These systems generally deal very well
with usual features of first-order programs, for example arrays, use
SMT solvers to discharge proof obligations and strive for the best
possible automation. We try to improve on this work by adding
higher-order features, trying to maintain the degree of automation and
ease of use.
\section{Conclusion}
\label{sec:conclusion}

\paragraph{Future Work.}

We are currently investigating the possibility to obtain first-order proof
obligations, to be able to use SMT solvers~\cite{RanTin-SMTLIB}, which would
avoid manual proofs in many cases. An obvious solution would be to reuse
existing encodings of higher-order logic in first-order logic, such as the one
by Pottier and Gauthier~\cite{pottier-gauthier-hosc}. This technique has been
successfully employed by the Pangolin system~\cite{regis-gianas-pottier-08}.
Another option comes from our observation that in many cases, the proof
obligations for higher-order functions can actually be expressed in
first-order logic, when one replaces top-level {\em quantifications} of
functional arguments and invariants by top-level {\em declarations}. Doing
this translation manually, we have obtained first-order proof
obligations for the {\tt Array.iter} function which are all discharged by
SMT solvers. It is future work to find out if such a transformation can always
be carried out.

Currently, our effect system is quite limited. One possible extension is the
addition of subtyping: an effectful function $f$ of type $τ ->^ετ'$ can be seen
as a function of type $τ ->^{ε'}τ'$, if $ε\subseteq ε'$. In particular, such a
subtyping operation should come at no cost for program verification; it should
still come for free that $f$ does not modify the variables in $ε'\setminus ε$. It
would also be interesting to see if an ``impure'' function of type $τ
->^\emptyset τ'$ can be seen as pure function $τ -> τ'$ (and be used in the
logic).

Another possibility would be to modify the effect system itself to allow for
aliasing. In the setting with the aliasing restriction, our regions actually
become {\em singleton regions}, regions containing exactly one memory
location. Other, more powerful effect systems allow for {\em group regions},
with the possibility to temporarily {\em focus } on a particular memory cell
in that region. We see no fundamental difficulty in adopting such an effect
system, but haven't worked out the details.

Finally, from a programmer's point of view, it would be nice to have more
features in the programming language, such as recursive types and algebraic
datatypes. As long as so defined types verify a {\em positivity condition},
%TODO citation needed
these extensions would be completely orthogonal to the weakest precondition
calculus and the type and effect system used.

\bibliographystyle{plain}
\bibliography{biblio}{}

\end{document}